---
title: 'US vs World: China/Hunan Virus 2019'
author: "David J Jackson"
Date: "3/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pracma)
library(plotly)
library(scales)
library(prophet)
library(readxl)
library(httr)
library(lubridate)
library(coronavirus)
library(RSQLite)
```
```{r, echo=FALSE}
rm(list=ls())
#these libraries are necessary
#create the URL where the dataset is stored with automatic upDates every day
db <- dbConnect(SQLite(), dbname="../db/CORVID.sqlite3")

JHUDATA<- dbGetQuery(db, "SELECT country,Date,confirmed,deaths,recovered FROM JHUDATA;")
colnames(JHUDATA) <- c("Countries","Date","Cases","Deaths","Recovered")
JHUDATA$Date <- as.Date(JHUDATA$Date)
dbDisconnect(db)

```

#### Data Date Range
```{r,echo=FALSE}
min(JHUDATA$Date)
max(JHUDATA$Date)
```

### The World: Summary by Day for Cases,Deaths, and Percent of Deaths

```{r,echo=FALSE}
# Calculate Summary Cases and Deaths(World)
# World
# JHUDATA Summary by Year
JHUDATA_Year <- JHUDATA %>% group_by(Yearly =floor_date(Date,"year")) %>%                                             summarise(
  
                                    tc = sum(Cases),
                                    td = sum(Deaths),
                                    mean_cases = mean(Cases),
                                    mean_deaths = mean(Deaths),
                                    death_rate = td/tc,
                                    Count = n())

# Summary by Month
JHUDATA_monthly <- JHUDATA %>% group_by(Monthly =floor_date(Date,"month")) %>%                                            summarise(
                                        tc = sum(Cases),
                                        td = sum(Deaths), 
                                        mean_cases = mean(Cases),
                                        mean_deaths = mean(Deaths),                                                        death_rate = td/tc,
                                        Count = n())

## Weekly Summary
JHUDATA_weekly <- JHUDATA %>% group_by(Weekly =floor_date(Date,"week")) %>%                                             summarise(
                                      tc = sum(Cases),
                                      td = sum(Deaths),
                                      mean_cases = mean(Cases),
                                      mean_deaths = mean(Deaths),
                                      death_rate = td/tc,
                                      Count = n())

## Daily Summary
JHUDATA_daily <- JHUDATA %>% group_by(Daily =floor_date(Date,"day")) %>%                                              summarise(
                                    tc = sum(Cases),
                                    td = sum(Deaths),  
                                    mean_cases = mean(Cases),
                                    mean_deaths = mean(Deaths),
                                    death_rate = td/tc,
                                    Count = n())
```
```{r, echo=FALSE}
JHUDATA_daily %>% plot_ly() %>% add_lines(x=~Daily,y=~tc) %>% 
        layout(title="The World: Total Cases By Day")


JHUDATA_daily %>% plot_ly() %>% add_lines(x=~Daily,y=~td) %>% 
        layout(title="The World: Total Deaths By Day")

JHUDATA_daily %>% plot_ly() %>% add_lines(x=~Daily,y=~death_rate) %>% 
        layout(title="The World: Percent of Cases Ending in Death by Week") 
      
```


```{r, echo=FALSE}
# By Country
# Weekly
Countries_yearly <- JHUDATA %>% group_by(Countries,Yearly =floor_date( Date,"year")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 death_rate = td/tc,
                                                              Count = n())

# Monthly
# Monthly
Countries_monthly <- JHUDATA %>% group_by(Countries,Monthly =floor_date( Date,"month")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 death_rate = td/tc,
                                                              Count = n())

# Weekly
Countries_weekly <- JHUDATA %>% group_by(Countries,Weekly =floor_date( Date,"week")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 death_rate = td/tc,
                                                              Count = n())
## Daily
Countries_daily <- JHUDATA %>% group_by(Date) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 death_rate = td/tc,
                                                              Count = n())
```


#### World minus US

```{r}
JHUDATA_minus_us <- Countries_weekly %>% filter(Countries != "US" )
```

# USA Only
```{r}
US_yearly <- Countries_yearly %>% filter(Countries == "US" )
US_monthly <- Countries_monthly %>% filter(Countries == "US" )
US_weekly <- Countries_weekly %>% filter(Countries == "US" )
```

### The World: Summary by Day for Cases,Deaths, and Percent of Deaths

```{r, echo=FALSE}
JHUDATA_daily %>% plot_ly() %>% add_lines(x=~Daily,y=~tc) %>% 
        layout(title="The World: Total Cases By Day")


JHUDATA_daily %>% plot_ly() %>% add_lines(x=~Daily,y=~td) %>% 
        layout(title="The World: Total Deaths By Day")

JHUDATA_daily %>% plot_ly() %>% add_lines(x=~Daily,y=~death_rate) %>% 
        layout(title="The World: Percent of Cases Ending in Death by Week") 
      
```

### US Summary by Week of Cases,Deaths and Percent of Deaths

```{r,echo=FALSE}

US_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~tc) %>% 
        layout(title="USA: Total Cases By Week")
US_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~td) %>% 
        layout(title="USA: Total Deaths By Week")
US_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~death_rate) %>% 
        layout(title="USA Percent of Cases Ending in Death by Week")
```

## US vs World: Percent of Death by Week

```{r,echo=FALSE}
plot_ly() %>% add_lines(data=US_weekly,x=~Weekly,y=~death_rate,name="USA") %>% 
        add_lines(data=JHUDATA_weekly,x=~Weekly,y=~death_rate,name="World") %>%
        layout(title="USA/World Percent of Cases Ending in Death by Week")
```

### The World minus US: Summary by Week of Cases,Deaths, and Percent of Deaths

```{r, echo=FALSE}

```

### Daly Totals form Cases,Deaths and Death Rate

```{r, echo=FALSE}
df <-JHUDATA %>% group_by(Date) %>% summarise(
                                          TotalCases=sum(Cases),
                                         TotalDeaths = sum(Deaths),
                                         DeathRate = TotalDeaths/TotalCases)
df %>% plot_ly() %>% add_bars(x=~Date,y=~TotalCases) %>% layout(title="Total Confrimed Cases by Day")

df %>% plot_ly() %>% add_bars(x=~Date,y=~TotalDeaths) %>% layout(title="Total Confrimed Deaths by Day")

df %>% plot_ly() %>% add_bars(x=~Date,y=~DeathRate) %>% layout(title="Total Confrimed Deaths by Day")
```

### US Daily Reports for Cases,Deaths and Death Rate

```{r,echo=FALSE}

df.us <-JHUDATA %>% filter(Countries =="US") %>% group_by(Date) %>% summarise(
                                          TotalCases=sum(Cases),
                                         TotalDeaths = sum(Deaths),
                                         DeathRate = TotalDeaths/TotalCases)

df.us %>% plot_ly() %>% add_bars(x=~Date,y=~TotalCases) %>% layout(title="US Total Confrimed Cases by Day")

df.us %>% plot_ly() %>% add_bars(x=~Date,y=~TotalDeaths) %>% layout(title="US Total Confrimed Deaths by Day")

df.us %>% plot_ly() %>% add_bars(x=~Date,y=~DeathRate) %>% layout(title="US Death Rate (as Percentage) by Day")
```