---
title: 'John Hopkins : China (novel-COVID-19) Flu'
author: "David J Jackson"
date: "2020-03-25 22:15 EST/EDT"
output: html_document
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(lubridate)
library(plotly)
library(RSQLite)
library(pracma)
library(forcats)
```
```{r}
rm(list=ls())
```
```{r,echo=FALSE}
db <- dbConnect(SQLite(), dbname="../db/CORVID.sqlite3")

JHU <- dbGetQuery(db, "SELECT country,date,confirmed,deaths,recovered FROM JHUDATA;")
colnames(JHU) <- c("Countries","Reported","Cases","Deaths","Recovered")
JHU$DeathRate <- JHU$Deaths/JHU$Cases 
JHU$Reported <- as.Date(JHU$Reported)

JHUDAILY <- JHU %>% group_by(Countries) %>% filter(Reported == max(Reported))

```
```{r}

db <- dbConnect(SQLite(), dbname="../db/CORVID.sqlite3")
JHUDAILY$date <- as.character(JHUDAILY$Reported)
dbWriteTable(db, "JHULASTEST",JHUDAILY ,overwrite=TRUE)  
dbListTables(db)
dbDisconnect(db)
```

## Data Source: John Hopkins Coronavirus Resource Center
* URL: https://coronavirus.jhu.edu/map.html
* GitHub: https://github.com/CSSEGISandData/COVID-19.git

#### Plot Total Cases by Countries

```{r,echo=FALSE}
JHU_Countries_reported <- JHU %>% group_by(Countries,Reported) %>%                                 summarise(
                       Cases = sum(Cases),
                       TotalDeaths = sum(Deaths),
                       DeathRate =     TotalDeaths/Cases,
                                      Count = n()) %>% arrange(-Cases)
                                     
```
```{r,echo=FALSE}
JHU_Countries <- JHU %>% group_by(Countries) %>%                                 summarise(
                       Cases = sum(Cases),
                       TotalDeaths = sum(Deaths),
                       DeathRate =     TotalDeaths/Cases,
                                      Count = n()) %>% arrange(-Cases)
                                     
```


#### Total Confimed  Cases by Countries

```{r,echo=FALSE}
JHU %>% filter(Cases <= 250) %>%
   ggplot() + geom_col(aes(x=reorder(Countries,Cases),y=Cases)) +
        labs(title ="Countries with Confirmed Cases <= 250") +
  coord_flip() + scale_y_continuous(labels = scales::comma) 
```
```{r,echo=FALSE}
JHU %>% filter(Cases >=125) %>%
  ggplot() + geom_col(aes(x=reorder(Countries,Cases),y=Cases)) +
        labs(title ="Countries with Confirmed Cases >= 125") +
  coord_flip() + scale_y_continuous(labels = scales::comma) 
```

```{r,echo=FALSE}
JHU %>% filter(Cases >=2000) %>%
  ggplot() + geom_col(aes(x=reorder(Countries,Cases),y=Cases)) +
        labs(title ="Countries with Confirmed Cases >= 2000") +
  coord_flip() + scale_y_continuous(labels = scales::comma) 
```
```{r, echo=FALSE}
JHU %>% filter(Cases >=1000) %>%
ggplot() + geom_col(aes(x=reorder(Countries,Cases),y=Cases)) +
        labs(title ="Countries with Confirmed Cases >=2500") +
  coord_flip() + scale_y_continuous(labels = scales::comma) 
```
```{r,echo=FALSE}
JHU  %>%
 ggplot() + geom_col(aes(x=reorder(Countries,Cases),y=Cases)) +
        labs(title ="XCountries with Confirmed Cases >= 8000") +
  coord_flip() + scale_y_continuous(labels = scales::comma) 
  
```
```{r,echo=FALSE}
 JHU %>% filter(Cases >=5000) %>%
ggplot() + geom_col(aes(x=reorder(Countries,Cases),y=Cases)) +
        labs(title ="Countries with Confirmed Cases >= 16000") +
  coord_flip() + scale_y_continuous(labels = scales::comma) 
  
```

#### Death Rate Percentage by Countries 

```{r, echo=FALSE}
 JHU %>% ggplot() + geom_col(aes(x=reorder(Countries,DeathRate),y=DeathRate)) +
        labs(title ="Death Rate Percentage: Countries with Confirmed Cases > 500",x="Percentage of Confirmed Case That End in Death") +
        scale_y_continuous(labels=scales::percent) + coord_flip()
```
```{r}
# 
JHU %>% ggplot() +  geom_col(aes(x=reorder(Countries,DeathRate),y=DeathRate)) +
        labs(title ="Death Rate Percentage:Countries with Confirmed Cases > 1000", 
        x="Percentage of Confirmed Case That End in Death") + 
        scale_y_continuous(labels=scales::percent) + coord_flip()
```
```{r}
#
JHU %>% ggplot() +geom_col(aes(x=reorder(Countries,DeathRate),y=DeathRate)) +
        labs(title ="Death Rate Percentage: Countries with Confirmed Cases > 5000", x="Percentage of Confirmed Case That End in Death") +              scale_y_continuous(labels=scales::percent) + coord_flip()

```

#### US Vs World: Total Cases,Total Deaths, Death Rate

```{r}
JHU_us <- JHU %>% filter(Countries=="US")
JHU_world <- JHU %>% filter(Countries !="US")
#
JHUDATA <- JHU_world %>% group_by(Reported) %>% summarise(
                                      Cases = sum(Cases),
                                      TotalDeaths = sum(Deaths),
                                      DeathRate = TotalDeaths/Cases,
                                      Count = n())
#
US_total_daily <- JHU_us %>% group_by(Reported) %>% summarise(
                                      Cases = sum(Cases),
                                      TotalDeaths = sum(Deaths),
                                      DeathRate = TotalDeaths/Cases,Count = n())

```

```{r, echo=FALSE}
ggplot() + geom_line(data=US_total_daily,aes(x=Reported,y=Cases,col="US")) +
  geom_line(data=JHUDATA,aes(x=Reported,y=Cases,col="World")) +
  scale_y_continuous(labels = scales::comma) +
  labs(title="US vs World: Total Confirmed Cases By Date")
```
```{r}
ggplot() + geom_line(data=US_total_daily,aes(x=Reported,y=TotalDeaths,col="US")) +
  geom_line(data=JHUDATA,aes(x=Reported,y=TotalDeaths,col="World")) +
  scale_y_continuous(labels = scales::comma) + 
  labs(title="US vs World: Total Deaths By Date")
```

```{r}
ggplot() + geom_line(data=US_total_daily,aes(x=Reported,y=DeathRate,col="US")) +
  geom_line(data=JHUDATA,aes(x=Reported,y=DeathRate,col="World")) +
  labs(title="US vs World: Death Rate as  Percentage By Date") +
  scale_y_continuous(labels=scales::percent, limits = c(0.0,0.4))
  
```
```{r}

JHUDAILY$E <- movavg(JHUDAILY$Cases, 12,"e")
JHUDAILY$ED <- JHUDAILY$Cases - JHUDAILY$E
ggplot(JHUDAILY) +geom_line(aes(x=Reported,y=Cases,col="Cases")) +
  geom_line(aes(x=Reported,y=E,col="Exponential")) + labs(title="Total Cases vs Moving Average",subtitle = "( 12 Day Exponential Moving Average)")
```
```{r}
JHUDAILY$E1 <- movavg(JHUDAILY$td, 12,"e")
ggplot(JHUDAILY) +geom_line(aes(x=Daily,y=td,col="Total Deaths")) +
  geom_line(aes(x=Daily,y=E1,col="Exponential")) + labs(title="Total Deaths vs Moving Average",subtitle ="( 12 Day Exponential Moving Average)" )
```
```{r}
JHUDAILY$E2 <- movavg(JHUDAILY$perdeaths, 12,"e")
ggplot(JHUDAILY) +geom_line(aes(x=Daily,y=perdeaths,col="Total Deaths")) +
  geom_line(aes(x=Daily,y=E2,col="Exponential")) + labs(title="Death Rate vs Moving Average",subtitle ="( 12 Day Exponential Moving Average)" )
```

###  Just the Facts

```{r}
world_less_us <- JHU%>% filter(Countries !="US")
sum(na.omit(world_less_us$Deaths))/sum(na.omit(world_less_us$Cases)) *100

sum(na.omit(JHU$Deaths))/sum(na.omit(JHU$Cases)) *100

US <-JHU %>% filter(Countries=="US")
sum(na.omit(US$Deaths))/sum(na.omit(US$Cases)) * 100


```
library(RSQLite)
db <- dbConnect(SQLite(), dbname="../db/CORVID.sqlite3")
dta$date <- as.character(dta$date)
dbWriteTable(db, "JHULASTEST",JHU ,overwrite=TRUE)    
    
 

                 